Our Future Nuclear Data Needs

### 2019-01-01 nuclear_data_needs/drafts/nuclear_data_needs_122018.tex
Adding old draft versions of review article

%Authors, affiliations address.
\author{Lee A. Bernstein$^{1,2}$, David A. Brown$^3$, \\Arjan J. Koning$^4$, Brad T. Rearden$^5$,  \\Catherine E. Romano$^6$, Alejandro A. Sonzogni$^3$, Andrew S. Voyles$^{1,2}$, and Walid  Younes$^7$
\affil{$^1$Nuclear Science Division, Lawrence Berkeley National Laboratory,  Berkeley, CA 94720, USA; email: labernstein@lbl.gov}
\affil{$^2$Department of Nuclear Engineering, University of California, Berkeley,  Berkeley, CA 94720, USA}
\affil{$^3$National Nuclear Data Center, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affil{$^4$Arjan's affiliations\textred{Arjan - please provide!}}
\affil{$^5$Brad's affiliations\textred{Brad - please provide!}, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^6$Reactor and Nuclear Systems Division, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^7$\textred{Add Walid's division here}, Lawrence Livermore National Laboratory,  Livermore, CA 94550, USA}}

@@

Low-energy nuclear science (LENS) \enquote{straddles the fence} between curiosity- and application-driven pursuits.  On the curiosity-driven side, low-energy ($E_x <$ 1--4 MeV) nuclear structure offers a unique laboratory to examine the interplay between single-particle and collective behavior and to explore the transition from quantum to continuum behavior in a mesoscopic setting. LENS also plays an important role in nuclear astrophysics, improving models of stellar energy generation and allowing isotopic abundances to be used to inform models of the astrophysical environments where heavy nuclei are formed.  On the applied side, a well-quantified knowledge of low-lying nuclear structure and reactions is needed to model energy generation and isotope production for medical and industrial uses and for the national security and counter-proliferation communities.

@@

In the case of nuclear reaction data, the process is markedly different.  The data from EXFOR is used to guide a physics-based model calculation through a variety of variance minimization procedures, resulting in best estimates of mean values and their uncertainties, including covariances, which are put into the Evaluated Nuclear Data File \cite{Chadwick2011}.  Most of the effort put into the production of ENDF is provided by applications-oriented users due to its importance to national security, international counter-proliferation and nuclear energy.  A consequence of this application focus is that the vast majority of the data in ENDF concerns neutron-induced reactions at either thermal (25 meV), fast fission or 14.1 MeV.  This has profound implications for isotope production since many of the nuclides are produced through charged-particle induced reactions.

@@

The conversion method is not fully free of issues, either.  To start, it is based on a single set of measurements; also, its most basic premise, the description of a level-to-level spectrum using a particular formalism, has not been tested, as highly precise measurements of electron spectra for fission products above the inverse beta decay cross section threshold have not been published.    As a result, while more precise than the summation method, it is nevertheless considerably less reliable.     A recent article \cite{Son17} reviewed the different Fermi functions available for the calculation of spectra and showed that a linear correction factor of 6\%/MeV could explain the anomaly.

Given the dependence of the antineutrino signal on both decay and fission product yield data, a series of sensitivity study-guided integral experiments, along the lines of critical assembly integral benchmarks, are needed to validate the underlying nuclear data would provide a strong constraint on the underlying data.  One approach would involve irradiating fissile and fissionable samples in facilities with well-understood neutron spectra and comparing the observed delayed gamma-rays to a modeled spectrum produced using the Fission Induced Electromagnetic Response (FIER) code developed at Berkeley \cite{Matthews2018}.  FIER convolves fission product yield distributions and decay data to predict the gamma-ray spectrum after neutron irradiation.  The first example of its use involving a comparison to the delayed gamma-ray spectrum from a \ce{^{235}U} foil irradiated in the Godiva critical assembly indicated several cases were the underlying fission yields or decay data required improvement.

\textred{We seem to have plenty of space - should we add the FIER plot back in?}

@@

  \item Improving our knowledge of the \ce{^{238}U}(n,n')cross section.   \ce{^{238}U}(n,n') is a limiting factor in the design of fast spectrum reactors such as those with metallic cores or chloride salts. The 40\% uncertainty in this cross section in ENDF/B-VII.1 propagates to an uncertainty of 1.5\%$\Delta k$ in the criticality state of a sodium fast reactor (i.e. at the 2$\sigma$ level, $k_{eff}$ could range from 0.97 to 1.03 based on this uncertainty prediction) \cite{touran2016sensitivities}.
  \item Improved knowledge of the \ce{^{35}Cl}(n,p) cross section.  Changes from the 2006 release of ENDF/B-VII.0 to the 2011 release of ENDF/B-VII.1 are causing a 2-3\%$\Delta k$ change in reactivity in molten salt reactors. This data change will also impact repository analysis in salts and there are no benchmark experiments available for validation. Recent results from activation using a DD-neutron source \cite{voy18a} indicate both resonant behavior in this cross section and an average value significantly smaller value for  $2.2 < E_n (MeV) < 2.7$ \cite{Bat18].  
  \item Graphite data for high-temperature reactors as well as the Transient Reactor Test Facility (TREAT) facility have realized a number of updates that cause dramatic changes in criticality. Updates from ENDF/B-VII.0 to ENDF/B-VII.1 lead to a 1\%$\Delta k$ change in reactivity for the High Temperature Engineering Test Reactor (HTTR), leading to the calculations better matching experimental values. Proposed updates to the graphite thermal scattering data for the ENDF/B-VIII.0 release have been demonstrated to cause a 2\%$\Delta k$ change in reactivity for the TREAT M8CAL benchmark \cite{hawarithermal}.
  \item Thermal neutron scattering in molten salts, such as FLiBe. There are no cross-section data that represent the effect of the molecular bonds, leading to a computational bias of unknown magnitude in all reactor design calculations. The molecular bonds in water vs. Polyethylene can lead to a computation bias of 1-2\%$\Delta k$.  In addition, this effect is temperature dependent and is not well studied or understood. A recently developed benchmark experiment evaluation of the Molten Salt Reactor Experiment which consisted of uranium fuel FLiBe molten salt flowing through graphite channels demonstrates a bias of over 1\%$\Delta k$ from the experimental value \cite{shenzero}, with other presented results showing biases over 3\%$\Delta k$

@@

  \item Uncertainties provided with ENDF in the form of cross section covariance data are often not mature and may or may not represent the true uncertainty in the nuclear data. For the recent ENDF/B-VIII.0 release, the readme file provided by the National Nuclear Data Center (NNDC) provides a disclaimer warning against the use of covariance data without applying a sensitivity/uncertainty approach to adjust the uncertainties for application specific analysis using relevant benchmark experiments. This specialized process is generally not available to users and presents many complicated aspects even for experts in the field, leaving a large gap between data provided by nuclear data evaluators and those useful for application in safety and design calculations.
  \item Nuclear data are gathered from many different sources, not only ENDF but also the ENSDF as well as the Joint European Fission Fusion (JEFF) library, which often have differing representations of the same data. The process of sorting through the data to correct inconsistencies and benchmark the consolidated library against benchmark experiments is complex.
  \item Uncertainty data are not available or are incomplete for fission product yields, decay constants, and angular distributions, making it difficult to have high confidence in predicting the performance of advanced systems.

@@

%Disclosure
\section*{DISCLOSURE STATEMENT}
\textred{IThe authors are not aware of any affiliations, memberships, funding, or financial holdings that might be perceived as affecting the objectivity of this review. }

@@

Work at Lawrence Berkeley National Laboratory has been carried out  under the auspices of the U.S. Department of Energy by  Lawrence Berkeley National Laboratory and the U.S. Nuclear Data Program under contract \# DE-AC02-05CH11231. This work used the Savio computational cluster resource provided by the Berkeley Research Computing program at the University of California, Berkeley (supported by the UC Berkeley Chancellor, Vice Chancellor for Research, and Chief Information Officer).

@@

\bibliography{../../library}

>>>>

%Authors, affiliations address.
\author{Lee A. Bernstein$^{1,2}$, David A. Brown$^3$, \\Arjan J. Koning$^4$, Bradley T. Rearden$^5$,  \\Catherine E. Romano$^5$, Alejandro A. Sonzogni$^3$, Andrew S. Voyles$^{2}$, and Walid  Younes$^6$
\affil{$^1$Nuclear Science Division, Lawrence Berkeley National Laboratory,  Berkeley, CA 94720, USA; email: labernstein@lbl.gov}
\affil{$^2$Department of Nuclear Engineering, University of California, Berkeley,  Berkeley, CA 94720, USA}
\affil{$^3$National Nuclear Data Center, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affil{$^4$Nuclear Data Section, International Atomic Energy Agency, 1400 Vienna, Austria}
% \affil{$^5$Brad's affiliations\textred{Brad - please provide!}, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^5$Reactor and Nuclear Systems Division, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^6$Lawrence Livermore National Laboratory L-414, Livermore, CA 94551, USA}}

@@

Low-energy nuclear science (LENS) \enquote{straddles the fence} between curiosity- and application-driven pursuits.
 On the curiosity-driven side, low-energy nuclear structure offers a unique laboratory to examine the interplay between single-particle and collective behavior and to explore the transition from quantum to continuum behavior in a mesoscopic setting.
LENS also plays an important role in nuclear astrophysics, improving models of stellar energy generation and allowing isotopic abundances to be used to inform models of the astrophysical environments where heavy nuclei are formed.
 On the applied side, a well-quantified knowledge of low-lying nuclear structure and reactions is needed to model energy generation and isotope production for medical and industrial uses and for the national security and counter-proliferation communities.

@@

In the case of nuclear reaction data, the process is markedly different.
 The data from EXFOR is used to guide a physics-based model calculation through a variety of variance minimization procedures, resulting in best estimates of mean values and their uncertainties, including covariances, which are put into the one or more regional libraries such as the recently release of the Evaluated Nuclear Data File (ENDF), ENDF/B-VIII.0 \cite{endf8_0}.
 Most of the effort put into the production of ENDF is provided by applications-oriented users due to its importance to national security, international counter-proliferation and nuclear energy.
 A consequence of this application focus is that the vast majority of the data in ENDF concerns neutron-induced reactions at either thermal (25 meV), fast fission or 14.1 MeV.
 This has profound implications for isotope production since many of the nuclides are produced through charged-particle induced reactions.

@@

The conversion method is not fully free of issues, either.
 To start, it is based on a single set of measurements; also, its most basic premise, the description of a level-to-level spectrum using a particular formalism, has not been tested, as highly precise measurements of electron spectra for fission products above the inverse beta decay (IBD) cross section threshold have not been published.
   As a result, while more precise than the summation method, it is nevertheless considerably less reliable.
    A recent article \cite{Son17} reviewed the different Fermi functions available for the calculation of spectra and showed that a linear correction factor of 6\%/MeV could explain the anomaly.

\begin{figure}[b]
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.7\linewidth]{667_before_v2.png}

 % variation_curve.pdf: 612x792 pixel, 72dpi, 21.59x27.94 cm, bb=0 0 612 792
 \caption{Difference in the FIER-predicted and observed yield of \ce{^{132}I} as a function of time for a \ce{^{235}U} sample irradiated at the GODIVA critical assembly taken from \cite{Matthews2018}.  }
 \label{fig:fier_plot}
\end{figure}

Given the dependence of the antineutrino signal on both decay and fission product yield data, a series of sensitivity study-guided integral experiments, along the lines of critical assembly integral benchmarks, are needed to validate the underlying nuclear data would provide a strong constraint on the underlying data.
 One approach would involve irradiating fissile and fissionable samples in facilities with well-understood neutron spectra and comparing the observed delayed gamma-rays to a modeled spectrum produced using the Fission Induced Electromagnetic Response (FIER) code developed at Berkeley \cite{Matthews2018}.
 FIER convolves fission product yield distributions and decay data to predict the gamma-ray spectrum after neutron irradiation.
 The first example of its use involving a comparison to the delayed gamma-ray spectrum from a \ce{^{235}U} foil irradiated in the Godiva critical assembly indicated several cases were the underlying fission yields or decay data required improvement.
One example, where the yield of the \ce{^{132}I} needed to be adjusted upward by 0.24$\sigma$, is shown in \autoref{fig:fier_plot}.

@@

  \item Improving our knowledge of the \ce{^{238}U}(n,n')cross section, as discussed in \autoref{sec:n_transport} previously.
  \ce{^{238}U}(n,n') is a limiting factor in the design of fast spectrum reactors such as those with metallic cores or chloride salts.
The 40\% uncertainty in this cross section in ENDF/B-VII.1 propagates to an uncertainty of 1.5\%$\Delta k$ in the criticality state of a sodium fast reactor (i.e. at the 2$\sigma$ level, $k_{eff}$ could range from 0.97 to 1.03 based on this uncertainty prediction) \cite{touran2016sensitivities}.
  \item Improved knowledge of the \ce{^{35}Cl}(n,p) cross section.
 Changes from the 2006 release of ENDF/B-VII.0 to the 2011 release of ENDF/B-VII.1 are causing a 2-3\%$\Delta k$ change in reactivity in molten salt reactors.
This data change will also impact repository analysis in salts and there are no benchmark experiments available for validation.
Recent results from activation using a DD-neutron source \cite{Voyles2017} indicate both resonant behavior in this cross section and an average value significantly smaller value for  $2.2 < E_n \text{(MeV)} < 2.7$ \cite{Bat18}.  
  \item Graphite data for high-temperature reactors as well as the Transient Reactor Test Facility (TREAT) facility have realized a number of updates that cause dramatic changes in criticality.
Updates from ENDF/B-VII.0 to ENDF/B-VII.1 lead to a 1\%$\Delta k$ change in reactivity for the High Temperature Engineering Test Reactor (HTTR), leading to the calculations better matching experimental values.
Proposed updates to the graphite thermal scattering data for the ENDF/B-VIII.0 release have been demonstrated to cause a 2\%$\Delta k$ change in reactivity for the TREAT M8CAL benchmark \cite{hawarithermal}.
  \item Thermal neutron scattering in molten salts, such as FLiBe.
There are no cross-section data that represent the effect of the molecular bonds, leading to a computational bias of unknown magnitude in all reactor design calculations.
The molecular bonds in water vs. Polyethylene can lead to a computation bias of 1-2\%$\Delta k$.
 In addition, this effect is temperature dependent and is not well studied or understood.
A recently developed benchmark experiment evaluation of the Molten Salt Reactor Experiment which consisted of uranium fuel FLiBe molten salt flowing through graphite channels demonstrates a bias of over 1\%$\Delta k$ from the experimental value \cite{shenzero}, with other presented results showing biases over 3\%$\Delta k$

@@

  \item Uncertainties provided with ENDF in the form of cross section covariance data are often not mature and may or may not represent the true uncertainty in the nuclear data.
For the recent ENDF/B-VIII.0 release, the readme file provided by the National Nuclear Data Center (NNDC) provides a disclaimer warning against the use of covariance data without applying a sensitivity/uncertainty approach to adjust the uncertainties for application specific analysis using relevant benchmark experiments.
This specialized process is generally not available to users and presents many complicated aspects even for experts in the field, leaving a large gap between data provided by nuclear data evaluators and those useful for application in safety and design calculations.
  \item Nuclear data are gathered from many different sources, not only ENDF but also the ENSDF as well as the Joint European Fission Fusion (JEFF) library, which often have differing representations of the same data.
The process of sorting through the data to correct inconsistencies and benchmark the consolidated library against benchmark experiments is complex.
  \item Uncertainty data are not available or are incomplete for fission product yields, decay constants, and angular distributions, making it difficult to have high confidence in predicting the performance of advanced systems.  In the case of fission product yields, there is not even a format for covariance data in the ENDF format, harkening back to the admonition in  \autoref{sec:pipeline}.

@@

%Disclosure
\section*{DISCLOSURE STATEMENT}
The authors are not aware of any affiliations, memberships, funding, or financial holdings that
might be perceived as affecting the objectivity of this review.

@@

The work by staff from several laboratories was performed under the auspices of the US Department of Energy at Oak Ridge National Laboratory under Contract DE-AC0500OR22725,  Los Alamos National Laboratory under Contract DE-AC52-06NA25396, Lawrence Livermore National Laboratory under Contract DE- AC52-07NA27344, Pacific Northwest National Laboratory under contract DE-AC05-76RLO1830, Lawrence Berkeley National Laboratory under Contract DE-AC02-05CH11231, and Brookhaven National Laboratory under Contract DE-AC02-98CH10886. In addition, this work was funded in part by the Consortium for Verification Technology under Department of Energy National Nuclear Security Administration award number DE-NA0002534.

A portion of this work used the Savio computational cluster resource provided by the Berkeley Research Computing program at the University of California, Berkeley (supported by the UC Berkeley Chancellor, Vice Chancellor for Research, and Chief Information Officer).

@@

\bibliography{/Users/nssc/library.bib}

### 2019-01-01 nuclear_data_needs/drafts/nuclear_data_needs_dbrown.tex
Adding old draft versions of review article

%Authors, affiliations address.
\author{Lee A. Bernstein$^{1,2}$, David A. Brown$^3$, \\Arjan J. Koning$^4$, Bradley T. Rearden$^5$,  \\Catherine E. Romano$^5$, Alejandro A. Sonzogni$^3$, Andrew S. Voyles$^{2}$, and Walid  Younes$^6$
\affil{$^1$Nuclear Science Division, Lawrence Berkeley National Laboratory,  Berkeley, CA 94720, USA; email: labernstein@lbl.gov}
\affil{$^2$Department of Nuclear Engineering, University of California, Berkeley,  Berkeley, CA 94720, USA}
\affil{$^3$National Nuclear Data Center, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affil{$^4$Nuclear Data Section, International Atomic Energy Agency, 1400 Vienna, Austria}
% \affil{$^5$Brad's affiliations\textred{Brad - please provide!}, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^5$Reactor and Nuclear Systems Division, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^6$Lawrence Livermore National Laboratory L-414, Livermore, CA 94551, USA}}

@@

Low-energy nuclear science (LENS) \enquote{straddles the fence} between curiosity- and application-driven pursuits.
 On the curiosity-driven side, low-energy nuclear structure offers a unique laboratory to examine the interplay between single-particle and collective behavior and to explore the transition from quantum to continuum behavior in a mesoscopic setting.
LENS also plays an important role in nuclear astrophysics, improving models of stellar energy generation and allowing isotopic abundances to be used to inform models of the astrophysical environments where heavy nuclei are formed.
 On the applied side, a well-quantified knowledge of low-lying nuclear structure and reactions is needed to model energy generation and isotope production for medical and industrial uses and for the national security and counter-proliferation communities.

@@

\begin{figure}[b]
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.7\linewidth]{667_before_v2.png}

@@

  \item Uncertainty data are not available or are incomplete for fission product yields, decay constants, and angular distributions, making it difficult to have high confidence in predicting the performance of advanced systems.  In the case of fission product yields, there is not even a format for covariance data in the ENDF format, harkening back to the admonition in  \autoref{sec:pipeline}.

@@

The work by staff from several laboratories was performed under the auspices of the US Department of Energy at Oak Ridge National Laboratory under Contract DE-AC0500OR22725,  Los Alamos National Laboratory under Contract DE-AC52-06NA25396, Lawrence Livermore National Laboratory under Contract DE- AC52-07NA27344, Pacific Northwest National Laboratory under contract DE-AC05-76RLO1830, Lawrence Berkeley National Laboratory under Contract DE-AC02-05CH11231, and Brookhaven National Laboratory under Contract DE-AC02-98CH10886. In addition, this work was funded in part by the Consortium for Verification Technology under Department of Energy National Nuclear Security Administration award number DE-NA0002534.

A portion of this work used the Savio computational cluster resource provided by the Berkeley Research Computing program at the University of California, Berkeley (supported by the UC Berkeley Chancellor, Vice Chancellor for Research, and Chief Information Officer).

@@

\bibliography{/Users/nssc/library.bib}

>>>>

%Authors, affiliations address.
\author{Lee A. Bernstein$^{1,2}$, David A. Brown$^3$, \\Arjan J. Koning$^4$, Brad T. Rearden$^5$,  \\Catherine E. Romano$^5$, Alejandro A. Sonzogni$^3$, Andrew S. Voyles$^{1,2}$, and Walid  Younes$^6$
\affil{$^1$Nuclear Science Division, Lawrence Berkeley National Laboratory,  Berkeley, CA 94720, USA; email: labernstein@lbl.gov}
\affil{$^2$Department of Nuclear Engineering, University of California, Berkeley,  Berkeley, CA 94720, USA}
\affil{$^3$National Nuclear Data Center, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affil{$^4$Nuclear Data Section, International Atomic Energy Agency, 1400 Vienna, Austria}
% \affil{$^5$Brad's affiliations\textred{Brad - please provide!}, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^5$Reactor and Nuclear Systems Division, Oak Ridge National Laboratory, Oak Ridge, TN 37830, USA}
\affil{$^6$Lawrence Livermore National Laboratory L-414, Livermore, CA 94551, USA}}

@@

Low-energy nuclear science (LENS) \enquote{straddles the fence} between curiosity- and application-driven pursuits.
 On the curiosity-driven side, low-energy ($E_x <$ 1--4 MeV) \textred{DAB: This is very restrictive!} nuclear structure offers a unique laboratory to examine the interplay between single-particle and collective behavior and to explore the transition from quantum to continuum behavior in a mesoscopic setting.
LENS also plays an important role in nuclear astrophysics, improving models of stellar energy generation and allowing isotopic abundances to be used to inform models of the astrophysical environments where heavy nuclei are formed.
 On the applied side, a well-quantified knowledge of low-lying nuclear structure and reactions is needed to model energy generation and isotope production for medical and industrial uses and for the national security and counter-proliferation communities.

@@

\begin{figure}[b]
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.7\linewidth]{667_before_alt.png}

@@

  \item Uncertainty data are not available or are incomplete for fission product yields, decay constants, and angular distributions, making it difficult to have high confidence in predicting the performance of advanced systems.  \textred{In the case of fission product yields, there is not even a format for covariance data in the ENDF format, harkening back to the admonition in  \autoref{sec:pipeline}.}

@@

Work at Lawrence Berkeley National Laboratory has been carried out  under the auspices of the U.S. Department of Energy by  Lawrence Berkeley National Laboratory and the U.S. Nuclear Data Program under contract \# DE-AC02-05CH11231.
This work used the Savio computational cluster resource provided by the Berkeley Research Computing program at the University of California, Berkeley (supported by the UC Berkeley Chancellor, Vice Chancellor for Research, and Chief Information Officer).

This work was performed under the auspices of the U.S. Department of Energy by Lawrence Livermore National Security,
LLC, Lawrence Livermore National Laboratory under Contract \# DE-AC52-07NA27344.

@@

\bibliography{../../library}
